---
title: "Exercise 2.8.2"
author: "Geoff Hing"
date: "1/1/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library('qss')
data(gay, package = "qss")
```

# Changing Minds on Gay Marriage

## 1.

> Using the baseline interview wave before the treatment is administered, examine
whether randomization was properly conducted. Base your analysis on the
three groups of study 1: “same-sex marriage script by gay canvasser,” “same-sex
marriage script by straight canvasser” and “no contact.” Briefly comment on the
results.

```{r}
study.1 <- gay[gay$study == 1, ]

same.sex.gay <- study.1[study.1$treatment == 'Same-Sex Marriage Script by Gay Canvasser', ]
same.sex.straight <- study.1[study.1$treatment == 'Same-Sex Marriage Script by Straight Canvasser', ]
no.contact <- study.1[study.1$treatment == 'No Contact', ]

num.same.sex.gay <- dim(same.sex.gay)[1]
num.same.sex.straight <- dim(same.sex.straight)[1]
num.no.contact <- dim(no.contact)[1]
num.total <- dim(study.1)[1]

num.same.sex.gay / num.total
num.same.sex.straight / num.total
num.no.contact / num.total

num.same.sex.gay - num.same.sex.straight
(num.same.sex.gay - num.same.sex.straight) / num.total

```

In study 1, more than half of the subjects received no contact. 748 more subjects received the same-sex marriage script by a gay canvasser than by a straight canvassar, a little over one percent more.


## 2. Average treatment effects of gay and straight canvassers

> The second wave of the survey was implemented two months after canvassing.
Using study 1, estimate the average treatment effects of gay and straight canvassers
on support for same-sex marriage, separately. Give a brief interpretation
of the results.

```{r}
tapply(study.1$ssm, study.1$treatment, mean)
```

People who were contacted by canvassers using the same-sex marriage script were more supportive of same-sex marriage than those who had no contact with canvassers. Interestingly, people contacted by a gay canvassar with the recycling script had the highest support. People contact by a gay canvassar reading the sam-sex marriage script were only slightly more supportive of same-sex marriage than those contact by a straight canvasser with the same script.

## 3. Treatment and canvassers

> The study contained another treatment that involves contact, but does not
involve using the gay marriage script. Specifically, the authors used a script
to encourage people to recycle. What is the purpose of this treatment? Using
study 1 and wave 2, compare outcomes from the treatment “same-sex marriage
script by gay canvasser” to “recycling script by gay canvasser.” Repeat the same
for straight canvassers, comparing the treatment “same-sex marriage script by
straight canvasser” to “recycling script by straight canvasser.” What do these
comparisons reveal? Give a substantive interpretation of the results.

```{r}
study.1.wave.2 <- subset(study.1, wave == 2)
tapply(study.1.wave.2$ssm, study.1.wave.2$treatment, mean)
```

The authors used the recycling script to try to differentiate between responses to contact with someone who identifies as gay and responses to the script in particular.

Comparing the role of a gay canvasser with the same-sex marriage and recycling scripts, the impact of a gay canvassar reading the recycling script was more significant. People who got that the treatment of a gay canvassar were more supportive. With the same-sex marriage script, the impact of a gay canvasser was less profound. 

## 4.

> In study 1, the authors reinterviewed the respondents six different times (in
waves 2 to 7) after treatment, at two-month intervals. The last interview, in
wave 7, occurs one year after treatment. Do the effects of canvassing last? If
so, under what conditions? Answer these questions by separately computing the
average effects of straight and gay canvassers with the same-sex marriage script
for each of the subsequent waves (relative to the control condition).

```{r}
study.1.wave.3 <- subset(study.1, wave == 3)
study.1.wave.4 <- subset(study.1, wave == 4)
study.1.wave.5 <- subset(study.1, wave == 5)
study.1.wave.6 <- subset(study.1, wave == 6)
study.1.wave.7 <- subset(study.1, wave == 7)

# TODO: Figure out how to combine these all into one dataframe
tapply(study.1.wave.3$ssm, study.1.wave.3$treatment, mean)
tapply(study.1.wave.4$ssm, study.1.wave.4$treatment, mean)
tapply(study.1.wave.5$ssm, study.1.wave.5$treatment, mean)
tapply(study.1.wave.6$ssm, study.1.wave.6$treatment, mean)
tapply(study.1.wave.7$ssm, study.1.wave.7$treatment, mean)
```

It looks like, independent of treatment, support for same-sex marriage increased over the period of the study. The difference between support for those who were read the same-sex marriage script by a gay canvasser and a straight canvasser was greater after a year.


## 5.

> The researchers conducted a second study to replicate the core results of the first
study. In this study, same-sex marriage scripts are given only by gay canvassers.
For study 2, use the treatments “same-sex marriage script by gay canvasser” and
“no contact” to examine whether randomization was appropriately conducted.
Use the baseline support from wave 1 for this analysis.

```{r}
study.2 <- gay[gay$study == 2, ]

study.2.wave.3 <- subset(study.2, wave == 3)
study.2.wave.4 <- subset(study.2, wave == 4)
study.2.wave.5 <- subset(study.2, wave == 5)
study.2.wave.6 <- subset(study.2, wave == 6)
study.2.wave.7 <- subset(study.2, wave == 7)

# TODO: Figure out how to combine these all into one dataframe
tapply(study.2.wave.3$ssm, study.2.wave.3$treatment, mean)
tapply(study.2.wave.4$ssm, study.2.wave.4$treatment, mean)
tapply(study.2.wave.5$ssm, study.2.wave.5$treatment, mean)
tapply(study.2.wave.6$ssm, study.2.wave.6$treatment, mean)
tapply(study.2.wave.7$ssm, study.2.wave.7$treatment, mean)
```
```

## 6.

> For study 2, estimate the treatment effects of gay canvassing using data from
wave 2. Are the results consistent with those of study 1?

## 7.

>Using study 2, estimate the average effect of gay canvassing at each subsequent
wave and observe how it changes over time. Note that study 2 did not have a
fifth or sixth wave, but the seventh wave occurred one year after treatment, as in
study 1. Draw an overall conclusion from both study 1 and study 2.