---
title: "Exercise 2.8.1"
output: html_notebook
---



```{r, setup, include=FALSE}
library(qss)
```

```{r, load_star, include=FALSE}
data(STAR, package = "qss")
```

## 1.

```{r}
STAR$kinder <- factor(STAR$classtype)
levels(STAR$kinder) <- c("small", "regular", "regular with aid")

STAR$race <- factor(STAR$race)
levels(STAR$race) <- c("white", "black", "Asian", "Hispanic", "Native American", "others")
```

## 2.

Students who had small kindergarden class sizes had a higher average score on reading tests than students in the other two class types. There was a negligible difference between the average scores in classes with or without an aid. 

There was a negligible difference in average math scores between students who had small and regular kindergarden classes. Students who were in a regular class size with an aid had a slightly lower average class size.

```{r}
means_reading <- tapply(STAR$g4reading, STAR$kinder, mean, na.rm = TRUE)
means_math <- tapply(STAR$g4math, STAR$kinder, mean, na.rm = TRUE)

print(means_reading)

print(means_math)
```
However, taking into account the standard deviation, the effect of small class size on reading test scores is quite small.

See [What does effect size tell you? | Simply Psychology](https://www.simplypsychology.org/effect-size.html)
```{r}
sd_reading <- sd(STAR$g4reading, na.rm = TRUE)
sd_math <- sd(STAR$g4math, na.rm = TRUE)

unname((means_reading[1] - means_reading[2]) / sd_reading)
```

## 3.

The differences between high scores (66th percentile) and low scores (33rd percentile) are negligible between students with small and regular kindergarten classes, for both math and reading tests. 

```{r}
qs_small <- quantile(STAR$g4reading[STAR$kinder == "small"], c(0.33, 0.66), na.rm = TRUE)
qs_reg <- quantile(STAR$g4reading[STAR$kinder == "regular"], c(0.33, 0.66), na.rm = TRUE)

qs_small
qs_reg
```

```{r}
qs_small <- quantile(STAR$g4math[STAR$kinder == "small"], c(0.33, 0.66), na.rm = TRUE)
qs_reg <- quantile(STAR$g4math[STAR$kinder == "regular"], c(0.33, 0.66), na.rm = TRUE)

qs_small
qs_reg
```

## 4.

Students who were in a small kindergarten classroom made up the largest share of students who were in multiple years of small classrooms.

```{r}
kinder_yearssmall_xtab <- table(kinder = STAR$kinder, yearssmall = STAR$yearssmall)
prop.table(kinder_yearssmall_xtab)
```
Students who spent four years in small classrooms had the highest reading scores. Students with one year in a small classroom also had a higher test score than those who spent no years. However, there was a negligible effect for students who spent two or three years in a smaller classroom. In fact, students who spent only two years in a small classroom had lower scores than students who had never had a small classroom.

This makes me wonder if the real effect is from small kindergarten classes. Students seem to benefit from this, and benefit the most if they continue to have a small class size for all four years. But, perhaps having a small class only has a negligible effect if a child did not have a small kindergarten class.

```{r}
tapply(STAR$g4reading, STAR$yearssmall, mean, na.rm = TRUE)
tapply(STAR$g4reading, STAR$yearssmall, median, na.rm = TRUE)
```
Again, there doesn't seem to be an effect of small kindergarten classes on math test scores.

```{r}
tapply(STAR$g4math, STAR$yearssmall, mean, na.rm = TRUE)
tapply(STAR$g4math, STAR$yearssmall, median, na.rm = TRUE)
```
## 5.

Regardless of kindergarten class size, white students had higher average test scores than minority students.

```{r}
STAR$white_minority <- as.factor(ifelse(
  STAR$race == "white",
  "white",
  ifelse(
    STAR$race == "black" | STAR$race == "Hispanic",
    "minority",
    NA
  )
))
k_regular <- STAR[STAR$kinder == "regular" & !is.na(STAR$white_minority),]
tapply(k_regular$g4reading, k_regular$white_minority, mean, na.rm = TRUE)
tapply(k_regular$g4math, k_regular$white_minority, mean, na.rm = TRUE)
```
However, while both white and minority students who were in small kindergarten classes had higher average reading test scores than those of the same race who did not, minority students had a 10 point higher average reading score. The increase in average score for white students was smaller.

```{r}
k_small <-  STAR[STAR$kinder == "small" & !is.na(STAR$white_minority),]
tapply(k_small$g4reading, k_small$white_minority, mean, na.rm = TRUE)
tapply(k_small$g4math, k_small$white_minority, mean, na.rm = TRUE)
```
```{r}
table(STAR$white_minority, STAR$kinder)
```

## 6.

```{r}
table(STAR$kinder, STAR$hsgrad)
```

Kindergarten size seems to have a negligble effect on graduation rate. Around 83-84% of students graduated regardless of their kindergarten class size.

```{r}
tapply(STAR$hsgrad, STAR$kinder, mean, na.rm = TRUE)
```

Spending 4 years in small classes seemed to impact graduation rates. Nearly 88% of students who spent four years in small classes graduated, compared with 83% of those who spent three years in small classes or no years. The graduation rate between students who spent one or two years in small classes is smaller than either of the aforementioned categories.

```{r}
tapply(STAR$hsgrad, STAR$yearssmall, mean, na.rm = TRUE)
```

White students graduated at a higher rate than minority students.

```{r}
tapply(STAR$hsgrad, STAR$white_minority, mean, na.rm = TRUE)
```

White and minority students who were in small kindergarten classes did not have significantly different graduation rates than those who were not.

```{r}
tapply(k_small$hsgrad, k_small$white_minority, mean, na.rm = TRUE)
```

For both white and minority students, students in small classes for four years had the highest graduation rates of any group. However, the difference in graduation rates between students with no years in small classes and those with four years was greater minority students than for white ones.

```{r}
small_0_yrs <-  STAR[STAR$yearssmall == 0 & !is.na(STAR$white_minority),]
small_1_yrs <-  STAR[STAR$yearssmall == 1 & !is.na(STAR$white_minority),]
small_2_yrs <-  STAR[STAR$yearssmall == 2 & !is.na(STAR$white_minority),]
small_3_yrs <-  STAR[STAR$yearssmall == 3 & !is.na(STAR$white_minority),]
small_4_yrs <-  STAR[STAR$yearssmall == 4 & !is.na(STAR$white_minority),]
tapply(small_0_yrs$hsgrad, small_0_yrs$white_minority, mean, na.rm = TRUE)
tapply(small_1_yrs$hsgrad, small_1_yrs$white_minority, mean, na.rm = TRUE)
tapply(small_2_yrs$hsgrad, small_2_yrs$white_minority, mean, na.rm = TRUE)
tapply(small_3_yrs$hsgrad, small_3_yrs$white_minority, mean, na.rm = TRUE)
tapply(small_4_yrs$hsgrad, small_4_yrs$white_minority, mean, na.rm = TRUE)

```

